\textbf{ Spectrophotometric parallaxes with linear models:\\ Accurate distances for luminous red-giant stars }

### 2018-10-15 paper/sp.tex
todo

% To-do items
% -----------
% - need a snippet for Table 1.
% - check with WISE experts that we have the correct WISE photometry
% - HWR: Figure 7: Caption needs to say what arrows represent.
% - Make sure we incorporate / respond to Ness comments from Aug 13 email.
% - Cite and describe Eilers et al.

@@

Another validation of the results can be obtained by looking at known
clusters or spatially compact objects in the data.
In \figurename~\ref{fig:clusters}, we show the parallax distribution from \gaia\ and
the spectrophotometric-parallax distribution from our model for stars that are
astrometrically confirmed members of known stellar clusters.
These clusters span some range in metallicity and age, and therefore test
the accuracy of the method for different kinds of populations.
With the exception of \acronym{M71}, these figures 
indicate that both \gaia\ and the spectrophotometric parallaxes appear to be unbiased
for these clusters and the range in abundances and ages they represent.
The case of cluster \acronym{M71} is troubling, but after searching for trends with
housekeeping data (also see \figurename~\ref{fig:residuals}) we don't find any correlations
with parallax offsets.
\begin{figure}
~\hfill\includegraphics[width=0.7\textwidth]{clusters.pdf}\hfill~
\caption{The \gaia\ astrometric parallaxes $\gparallax_n$
  and spectrophotometric parallaxes $\sparallax_n$ for member stars for
  four stellar clusters in the \apogee\ observing footprint. The spectrophotometric
  parallaxes and their inverse-variance-weighted mean and the astrometric parallaxes
  and their inverse-variance-weighted mean
  are shown as vertical lines.
  Cluster members were identified to be close to the cluster in celestial coordinates
  and comoving in proper motion.\label{fig:clusters}}
\end{figure}

@@

This \documentname\ and project delivers greatly improved distance estimates for
luminous red-giant stars.
The purpose of all this is to make global maps of the Milky Way.
Just as a demonstration, we show kinematics and element abundances for stars as a function
of position in the Milky Way disk in \figurename~\ref{fig:disk}.
In this visualization, it is possible to see the rotation of the disk and the
kinematic center of the Galaxy; that might even provide a distance estimate to the Galactic
Center.
It also shows extremely strong chemical gradients.
\begin{figure}
\includegraphics[width=\textwidth]{map.pdf}
\caption{A map of the kinematics and element abundances in the Milky Way disk.
  This map shows the rotation of the Galaxy out to large radius, and also
  indicates the kinematic center of the disk rotation. Interpretation of
  the quantitative information in this plot requires care (see text).\label{fig:disk}}
\end{figure}
In a companion paper (\citealt{eilers}), we are using this information to estimate
the (asymmetric-drift-corrected) circular-velocity curve for the Milky Way.

>>>>

% To-do items
% -----------
% - need a snippet for Table 1.
% - check with WISE experts that we have the correct WISE photometry
% - ACE: What fraction of spectral pixels are (typically, or in each training set) zeroed out?

@@

Another validation of the results can be obtained by looking at known
clusters or spatially compact objects in the data.
In \figurename~\ref{fig:clusters}, we show the parallax distribution from \gaia\ and
the spectrophotometric-parallax distribution from our model for stars that are
astrometrically confirmed members of known stellar clusters.
The cluster members were found by hand, looking at sky and proper-motion distributions
centered on literature values.
In each case we are plotting only very securely identified cluster members; that
is, membership is conservative.
These clusters span some range in metallicity and age, and therefore test
the accuracy of the method for different kinds of populations.
With the exception of \acronym{M71}, these figures 
indicate that both \gaia\ and the spectrophotometric parallaxes appear to be unbiased
for these clusters and the range in abundances and ages they represent.
The case of cluster \acronym{M71} is troubling, but after searching for trends with
housekeeping data (also see \figurename~\ref{fig:residuals}) we don't find any correlations
with parallax offsets.
\begin{figure}
~\hfill\includegraphics[width=0.7\textwidth]{clusters.pdf}\hfill~
\caption{The \gaia\ astrometric parallaxes $\gparallax_n$
  and spectrophotometric parallaxes $\sparallax_n$ for member stars for
  four stellar clusters in the \apogee\ observing footprint. The spectrophotometric
  parallaxes and their inverse-variance-weighted mean and the astrometric parallaxes
  and their inverse-variance-weighted mean
  are shown as vertical lines.
  Cluster members were identified to be close to the cluster in celestial coordinates
  and comoving in proper motion.\label{fig:clusters}}
\end{figure}

@@

This \documentname\ and project delivers greatly improved distance estimates for
luminous red-giant stars.
The purpose of all this is to make global maps of the Milky Way.
Just as a demonstration, we show kinematics and element abundances for stars as a function
of position in the Milky Way disk in \figurename~\ref{fig:disk}.
In this visualization, it is possible to see the rotation of the disk and the
kinematic center of the Galaxy; that might even provide a distance estimate to the Galactic
Center.
It also shows extremely strong chemical gradients.
\begin{figure}
\includegraphics[width=\textwidth]{map.pdf}
\caption{A map of the kinematics and element abundances in the Milky Way disk.
  Each arrow represents the three-space Galactocentric velocity of each star
  in the sample, projected onto the Galactic disk plane.
  To make these velocities, raw \gaia\ proper motions and \apogee\ radial velocities
  were used (in a standard Galactocentric transformation), so they are
  not in any way noise-deconvolved.
  Each arrow has a color corresponding to the \apogee\ reported metallicity.
  This map shows the rotation of the Galaxy out to large radius, and also
  indicates the kinematic center of the disk rotation. Interpretation of
  the quantitative information in this plot requires care.\label{fig:disk}}
\end{figure}
In a companion paper (\citealt{eilers}), we are using this information to estimate
the (asymmetric-drift-corrected) circular-velocity curve for the Milky Way.
